First Exclusive Measurement of Deep Virtual Compton Scattering off $^4$He: Toward the 3D tomography of nuclei

### 2017-04-20 eg6-letters/coherent/dvcs_he4_prl.tex
implementing adhoc comments, part B

One must overcome two major experimental challenges in measuring coherent 
nuclear DVCS, $eA\rightarrow e^\prime A^\prime\gamma$. First, the cross section 
of coherent scattering is suppressed due to the nuclear form factor, and 
second, the recoil nucleus ($A^\prime$) must be detected to ensure coherence.  
Figure \ref{fig:diags} illustrates the handbag diagram for coherent DVCS on 
$^4$He. Similar to the proton case, at large virtual photon's 4-momentum square 
$Q^2=-(k-k')^{2}$ and small squared momentum transfer $t=(p-p')^{2}$ (in terms 
of the electron, $k(k')$, and $^4$He, $p(p')$, four-vectors), the DVCS handbag 
diagram can be factorized into two parts \cite{Freund_Collins,Ji_Osborne}. The 
hard part includes the photon-quark interaction and is calculable in 
perturbative QCD. The non-perturbative part is parametrized in terms of GPDs, 
which embed the partonic structure of the hadron. The GPDs depend on the three 
variables $x$, $\xi$, and $t$ as introduced in Fig.~\ref{fig:diags}. The 
parameter $\xi$ relates to the Bjorken variable $x_{B}$: $\xi\approx 
{{x_B}\over{2-x_B}}$, where $x_B=\frac{Q^2}{2M\nu}$, $\nu=k^0-k^{0\prime}$, and 
$M$ is the proton mass. The parameter $x$ is the quark's internal loop momentum 
fraction and cannot be accessed experimentally. We in fact measure 
Compton form factors (CFF) \cite{Guidal:2013rya}, which are
complex amplitudes defined as:

@@

Until now, the only available data on nuclear DVCS was from the HERMES 
experiment \cite{Ellinghaus:2002zw}, where coherence in the reaction was based 
only on kinematic cuts on the measured scattered electron and real photon. The 
measurement was performed on a large set of nuclei ($^4$He, N, Ne, Kr and Xe), 
but the mixing of the coherent and incoherent processes could affect the
measurement significantly \cite{Guzey:2003jh}. In this regard, direct detection 
of the low-energy recoil nucleus is the best way to guarantee that the nucleus 
remains intact. The CEBAF large acceptance spectrometer in Hall-B at 
Jefferson Lab has been previously used successfully for DVCS measurements on the nucleon
\cite{Girod:2007aa,Gavalian:2009,Seder:2015,Pisano:2015,Jo:2015ema}.

In order to measure DVCS on $^4$He with a fully exclusive final state, we built 
a specialized radial time projection chamber (RTPC) for the detection and 
identification of low-energy recoiling light nuclei. The $^4$He nucleus is an 
ideal experimental target in this regard, as it is light enough to be detected 
in such a setting, while it is subject to significant nuclear effects 
\cite{JSeely} and has rather high density. A helium target leads to another 
important advantage, as the number of GPDs defined for a hadron depends on its 
spin. The structure of a spin-zero nucleus, such as $^4$He, is parametrized by 
only one chiral even GPD ($H_{A}(x,\xi,t)$) at leading twist, while 4 GPDs 
arise in the nucleon case. This significantly simplifies the interpretation of 
the data and allows a model independent extraction of the $^4$He CFF 
($\mathcal{H}_{A}$) presented at the end of this letter.

@@

In our kinematics, the recoil $^4$He nuclei produced in DVCS have low momentum,
averaging \~300 MeV/c. CLAS cannot detect such low energy $\alpha$ 
particles, so in order to ensure the exclusivity of the measurement, we built a 
small and light RTPC to complement CLAS. The RTPC is a $20$ cm long cylinder 
with a diameter of $15$ cm, positioned directly around the $^4$He gaseous 
target. The target is a $25$-cm-long and $6$-mm-diameter Kapton straw with 
$27~\mu$m thick walls. The RTPC was calibrated specifically for the detection of 
$^4$He nuclei by utilizing elastic scattering $e^4$He$\to e^\prime$$^4$He with
a 1.2~GeV electron beam.

@@

To ensure the interaction occurs at the partonic level and the DVCS handbag 
diagram is dominant, we select events with $Q^{2}$ greater than 1 
GeV$^{2}/$c$^{2}$. Exclusivity of the coherent DVCS reaction is optimized by 
applying a set of cuts on the following kinematic variables: the co-planarity 
angle ($\Delta\phi$), \textit{i.e.}, the angle between the ($\gamma,\gamma^*$) 
and ($\gamma^*$,$^4$He$'$) planes, the missing energy, the missing mass 
squared, the missing transverse momentum of the $e'^4$He$'\gamma$ system, the 
missing mass squared of the $e'^4$He$'$ system, and the angle ($\theta$) 
between the measured photon and the missing momentum of the $e'^4$He$'$ system.

@@

\begin{figure}[tb]
\hspace{-0.45cm}
\includegraphics[width=8.0cm]{figs/Q2_xB_t_Coh.pdf}
\caption{Coherent DVCS event distributions for $Q^{2}$ as a function of 
$x_{B}$ (left) and $Q^{2}$ as a function of $-t$ (right) after the 
exclusivity cuts.}
\label{fig:kin-coverage}
\end{figure}

@@

The DVCS and well-known Bethe-Heitler (BH) processes, in which the real 
photon is emitted by the incoming or the outgoing lepton, have the same final 
state and are indistinguishable. The amplitude of electroproduction of real
photons includes a sum of the amplitudes of these two processes. The BH 
amplitude is defined by the target form factors, while the DVCS amplitude is 
a combination of the form factors and GPDs. For our kinematics, the cross 
section of real photon electroproduction is dominated by the BH 
contribution, while the DVCS contribution is very small. Its effect is however
enhanced in the observables sensitive to the interference term, {\it e.g.} 
BSA. The three terms entering the cross section calculation,
the square of the BH and DVCS amplitudes and their interference term, depend on the
azimuthal angle
$\phi$ between the $(e,e^\prime)$ and $(\gamma^*,^4$He$^\prime)$ planes, as
shown for the nucleon in Ref.~\cite{Belitsky:2001ns} and for the spin-zero targets
in Refs.~\cite{Kirchner:2003wt,Belitsky:2008bz}. Based on this work, the BSA 
of a spin-zero hadron can be expressed as
\begin{equation}
\begin{split}
&A_{LU}(\phi) = \\
&\frac{\alpha_{0}(\phi) \, \Im m(\mathcal{H}_{A})}
{\alpha_{1}(\phi) + \alpha_{2}(\phi) \, \Re e(\mathcal{H}_{A}) + \alpha_{3}(\phi) \, 
\big( \Re e(\mathcal{H}_{A})^{2} + \Im m(\mathcal{H}_{A})^{2} \big)},
\end{split}
\label{eq:A_LU-coh}
\end{equation}
where $\Im m(\mathcal{H}_{A})$ and $\Re e(\mathcal{H}_{A})$ are the imaginary 
and real parts of the $^4$He CFF $\mathcal{H}_{A}$ and depend on the GPD $H_A$.  
The explicit expression of the kinematic factors $\alpha_i$ are written in
Ref.~\cite{Belitsky:2008bz} and are functions of Fourier harmonics in the 
azimuthal angle $\phi$, the nuclear form factor $F_A(t)$, $Q^2$, $x_B$,
$x_{A} = \frac{x_{B}M_{N}}{M_{A}}$, $y=\frac{Q^{2}}{2x_{A}M_{A}E_{beam}}$, and 
$\epsilon = \frac{2x_{A}M_{A}}{\sqrt{Q^{2}}}$, where $M_{A}$($M_{N}$) is the $^4$He 
(nucleon) mass. Using the different $\sin(\phi)$ and $\cos(\phi)$ contributions 
in these factors, it is possible to unequivocally extract both imaginary and real 
parts of $\mathcal{H}_{A}$ from the BSA.

In this work, the azimuthal dependence of the BSA, $A_{LU}$, has been studied 
for a wide range of kinematics. We identified two main backgrounds to our 
measurement, accidental coincidences and exclusive coherent $\pi^0$ production.  
The accidentals have particles originating from different events, and we 
estimated their contribution to be 4.1\% of our sample. We evaluated this 
contribution by selecting events passing all our cuts but with an electron and 
helium originating from different vertices. The $\pi^0$ production can easily
be mistaken for DVCS when one of the two photons from the $\pi^0$ 
decay is produced at low energy in the laboratory frame and remains undetected.  
To estimate the effect of this background, we developed an event generator 
tuned on the experimental yield of exclusive $\pi^0$ measured by our 
experiment. We used this generator together with a GEANT3 simulation of our 
detector to estimate the ratio between the number of $\pi^0$ events where the 
two photons are detected and those that are misidentified as DVCS events. This 
ratio is then multiplied by the measured yield of exclusive $\pi^0$ events to 
correct the DVCS data.  Depending on the kinematics, we found contaminations of 
2 to 4\%. The study of systematic uncertainties showed that the main 
contributions come from the choice of the DVCS exclusivity cuts (8\%) and the 
large binning size (5.1\%). However, added quadratically, the total systematic 
uncertainty is about 10\%, which is significantly smaller than statistical 
uncertainties in all kinematical bins.

@@

We acknowledge the staff of the Accelerator and Physics Divisions at Jefferson 
Lab for making this experiment possible. This work is supported by the U.S.  
Department of Energy, Office of Science, Office of Nuclear Physics, under 
contract DE-AC02-06CH11357.

>>>>

One must overcome two major experimental challenges in measuring coherent 
nuclear DVCS, $eA\rightarrow e^\prime A^\prime\gamma$. First, the cross section 
of coherent scattering is suppressed due to the nuclear form factor, and 
second, the recoil nucleus ($A^\prime$) must be detected to ensure coherence.  
Figure \ref{fig:diags} illustrates the handbag diagram for coherent DVCS on 
$^4$He, where four-vectors of the electrons, photons, and $^4$He are denoted by
$k/k^\prime$, $q/q^\prime$, and $p/p^\prime$ respectively. 
For large virtual photon's 4-momentum squared, $Q^2=-(k-k')^{2}$, and small squared momentum transfer, $t=(p-p')^{2}$, the DVCS handbag 
diagram can be factorized into two parts \cite{Freund_Collins,Ji_Osborne}. The 
hard part includes the photon-quark interaction and is calculable in 
perturbative QCD. The non-perturbative part is parametrized in terms of GPDs, 
which embed the partonic structure of the hadron. The GPDs depend on the three 
variables $x$, $\xi$, and $t$ as introduced in Fig.~\ref{fig:diags}. The 
parameter $\xi$ relates to the Bjorken variable $x_{B}$: $\xi\approx 
{{x_B}\over{2-x_B}}$, where $x_B=\frac{Q^2}{2M\nu}$, $\nu=k^0-k^{0\prime}$, and 
$M$ is the proton mass. The parameter $x$ is the quark's internal loop momentum 
fraction and cannot be accessed experimentally. We in fact measure 
Compton form factors (CFF) \cite{Guidal:2013rya}, which are
complex amplitudes defined as:

@@

Until now, the only available data on nuclear DVCS was from the HERMES 
experiment \cite{Ellinghaus:2002zw}, where coherence in the reaction was based 
only on kinematic cuts on the measured scattered electron and real photon. The 
measurement was performed on a large set of nuclei ($^4$He, N, Ne, Kr and Xe), 
but contamination from incoherent processes could influence the
measurement significantly \cite{Guzey:2003jh}.  Accordingly, direct detection 
of the low-energy recoil nucleus is critical in guaranteeing the nucleus 
remains intact.

The $^4$He nucleus is an 
ideal experimental target in this regard, as it is light enough to be detected 
in such a setting, while it is subject to significant nuclear effects 
\cite{JSeely} and has rather high density. A helium target leads to another 
important advantage, as the number of GPDs defined for a hadron depends on its 
spin. The structure of a spin-zero nucleus, such as $^4$He, is parametrized by 
only one chiral even GPD ($H_{A}(x,\xi,t)$) at leading twist, while 4 GPDs 
arise in the nucleon case. This significantly simplifies the interpretation of 
the data and allows a model independent extraction of the $^4$He CFF 
($\mathcal{H}_{A}$) presented at the end of this letter.

The CEBAF large acceptance spectrometer in Hall-B at 
Jefferson Lab has been previously used successfully for DVCS measurements on the nucleon
\cite{Girod:2007aa,Gavalian:2009,Seder:2015,Pisano:2015,Jo:2015ema}.

In order to measure DVCS on $^4$He with a fully exclusive final state, we built 
a specialized radial time projection chamber (RTPC) for the detection and 
identification of low-energy recoiling light nuclei.

@@

In our kinematics, the recoil $^4$He nuclei produced in DVCS have low momentum,
averaging 300 MeV/c. CLAS cannot detect such low energy $\alpha$ 
particles, so in order to ensure the exclusivity of the measurement, we built a 
small and light RTPC to complement CLAS. The RTPC is a $20$-cm-long cylinder 
with a diameter of $15$ cm, positioned directly around the $^4$He gaseous 
target. The target is a $25$-cm-long and $6$-mm-diameter Kapton straw with 
$27$-$\mu$m-thick walls. The RTPC was calibrated specifically for the detection of 
$^4$He nuclei by utilizing elastic scattering $e^4$He$\to e^\prime$$^4$He with
a 1.2~GeV electron beam.

@@

To ensure the interaction occurs at the partonic level and the DVCS handbag 
diagram is dominant, we select events with $Q^{2}$ greater than 1 
GeV$^{2}/$c$^{2}$. Exclusivity of the coherent DVCS reaction is optimized by 
applying a set of cuts on the following kinematic variables: the coplanarity 
angle ($\Delta\phi$) between the ($\gamma,\gamma^*$) 
and ($\gamma^*$,$^4$He$'$) planes, the missing energy, mass, and transverse
momentum of the $e'^4$He$'\gamma$ system, the missing mass squared of the $e'^4$He$'$ system, and the angle ($\theta$) 
between the measured photon and the missing momentum of the $e'^4$He$'$ system.

@@

\begin{figure}[tb]
\hspace{-0.45cm}
\includegraphics[width=8.5cm]{figs/Q2_xB_t_Coh.pdf}
\caption{Coherent DVCS event distributions for $Q^{2}$ as a function of 
$x_{B}$ (left) and $Q^{2}$ as a function of $-t$ (right) after the 
exclusivity cuts.}
\label{fig:kin-coverage}
\end{figure}

@@

The DVCS and well-known Bethe-Heitler (BH) processes, in which the real 
photon is emitted by the incoming or the outgoing lepton, have the same final 
state and are indistinguishable. The amplitude of electroproduction of real
photons includes a sum of the amplitudes of these two processes. The BH 
amplitude is defined by the target form factors, while the DVCS amplitude is 
a combination of the form factors and GPDs. For our kinematics, the cross 
section of real photon electroproduction is dominated by the BH 
contribution, while the DVCS contribution is very small. Its effect is however
enhanced in the observables sensitive to the interference term, {\it e.g.} 
BSA. The three terms entering the cross section calculation,
the square of the BH and DVCS amplitudes and their interference term, depend on the
azimuthal angle
$\phi$ between the $(e,e^\prime)$ and $(\gamma^*,^4$He$^\prime)$ planes, as
shown for the nucleon in Ref.~\cite{Belitsky:2001ns} and for the spin-zero targets
in Refs.~\cite{Kirchner:2003wt,Belitsky:2008bz}. Based on this work, the BSA 
of a spin-zero hadron can be expressed as
\begin{equation}
\begin{split}
&A_{LU}(\phi) = \\
&\frac{\alpha_{0}(\phi) \, \Im m(\mathcal{H}_{A})}
{\alpha_{1}(\phi) + \alpha_{2}(\phi) \, \Re e(\mathcal{H}_{A}) + \alpha_{3}(\phi) \, 
\big( \Re e(\mathcal{H}_{A})^{2} + \Im m(\mathcal{H}_{A})^{2} \big)},
\end{split}
\label{eq:A_LU-coh}
\end{equation}
where $\Im m(\mathcal{H}_{A})$ and $\Re e(\mathcal{H}_{A})$ are the imaginary 
and real parts of the $^4$He CFF $\mathcal{H}_{A}$ and depend on the GPD $H_A$.  
Explicit expressions of the purely kinematic factors $\alpha_i$ are written in
Ref.~\cite{Belitsky:2008bz} and are functions of Fourier harmonics in the 
azimuthal angle $\phi$, the nuclear form factor $F_A(t)$, $Q^2$, $x_B$,
$x_{A} = \frac{x_{B}M_{N}}{M_{A}}$, $y=\frac{Q^{2}}{2x_{A}M_{A}E_{beam}}$, and 
$\epsilon = \frac{2x_{A}M_{A}}{\sqrt{Q^{2}}}$, where $M_{A}$($M_{N}$) is the $^4$He 
(nucleon) mass. Using the different $\sin(\phi)$ and $\cos(\phi)$ contributions 
in these factors, it is possible to unequivocally extract both imaginary and real 
parts of $\mathcal{H}_{A}$ from the BSA.

In this work, the azimuthal dependence of the beam-spin asymmetry has been studied 
for a wide range of kinematics ($Q^2$ between 1 and 2.2 GeV$^2$/c$^2$, $x_B$ between 0.1 and 0.3, and $-t$ between 0.07 and 0.16 GeV$^2$/c$^2$). We identified two main backgrounds to our 
measurement, accidental coincidences and exclusive coherent $\pi^0$ production.  
The accidentals have particles originating from different events, and we 
estimated their contribution to be 4.1\% of our sample. We evaluated this 
contribution by selecting events passing all our cuts but with an electron and 
helium originating from different vertices. The $\pi^0$ production can easily
be mistaken for DVCS when one of the two photons from the $\pi^0$ 
decay is produced at low energy in the laboratory frame and remains undetected.  
To estimate the effect of this background, we developed an event generator 
tuned on the experimental yield of exclusive $\pi^0$ measured by our 
experiment. We used this generator together with a GEANT3 simulation of our 
detector to estimate the ratio between the number of $\pi^0$ events where the 
two photons are detected and those that are misidentified as DVCS events. This 
ratio is then multiplied by the measured yield of exclusive $\pi^0$ events to 
correct the DVCS data.  Depending on the kinematics, we found contaminations of 
2 to 4\%. The study of systematic uncertainties showed that the main 
contributions come from the choice of the DVCS exclusivity cuts (8\%) and the 
large binning size (5.1\%), where the values are relative and for $A_{LU}$ at $\phi=90^\circ$.
However, added quadratically, the total systematic uncertainty is about 10\%,
which is significantly smaller than statistical  uncertainties in all kinematical bins.

@@

We acknowledge the staff of the Accelerator and Physics Divisions at Jefferson 
Lab for making this experiment possible. This work is supported by the U.S.  
Department of Energy, Office of Science, Office of Nuclear Physics, under 
contracts DE-AC02-06CH11357 and DE-AC05-06OR23177.
